Computational biology in the 21st century:\\Scaling with compressive algorithms

### 2015-09-21 main.tex
cite de bruijn

(16S, shaped seeds, blocked
reactions, chemogenomics, de Bruijn)
\textbf{Definitions}

@@

\textbf{16S}

\textbf{de Bruijn graph}

@@

Given DNA or RNA reads from NGS technologies, the first task is to 
\emph{assemble} those fragments of sequence into contiguous sequences.
The assembly problem is analogous to the problem of 
reconstructing a book with all its pages torn out.
\emph{de novo} assembly is beyond the scope of this Review, but is
possible because the sequence is covered by many
overlapping reads~\cite{berger2013computational}; for this task, the \emph{de Bruijn graph}
data structure is commonly used.
Often, however, a \emph{reference genome} (or in the case of RNA, 
\emph{transcriptome}) is available for the organism being sequenced; the
establishment of a human reference genome was indeed the purpose of the Human
Genome Project.

@@

Techniques for reference-based read mapping typically rely on algorithmic 
approaches such as the Burrows-Wheeler transform (BWT)~\cite{berger2013computational}, 
which provides efficient string 
compression through a reversible transformation, while the FM-index data 
structure~\cite{berger2013computational} is a compressed substring index, 
based on the BWT, which provides efficient storage as well as fast search.
BWA (Burrows-Wheeler Aligner)~\cite{berger2013computational} uses the BWT, while the 
Bowtie2~\cite{berger2013computational} mapper further relies on the FM-index for 
efficient mapping of NGS reads.
The Genome Multitool (GEM) mapper~\cite{marco2012gem} also uses an FM-index 
coupled with dynamic programming in a compressed representation of the 
reference genome, in order to prune the search space 
when mapping reads to a reference genome.
Masai~\cite{siragusa2013fast} and mrsFAST \cite{berger2013computational} use an `approximate seed' approach to indexing
the space of possible matches, likewise pruning the search space; however, the
bulk of its runtime is spent on the extend phase.
State-of-the-art mapper mrsFAST-Ultra achieves improvements in efficiency based on machine architecture rather than leveraging redundancy in the data itself with near-perfect sensitivity, but
only for the case where there are no insertions and deletions (indels) \cite{hach2014mrsfast}.
Even with these approaches, read mapping remains a significant bottlenecks in
genomic research~\cite{berger2013computational}.

>>>>

\textbf{Definitions}

@@

@@

Given DNA or RNA reads from NGS technologies, the first task is to 
\emph{assemble} those fragments of sequence into contiguous sequences.
The assembly problem is analogous to the problem of 
reconstructing a book with all its pages torn out.
\emph{de novo} assembly is beyond the scope of this Review, but is
possible because the sequence is covered by many
overlapping reads~\cite{berger2013computational}; for this task, the \emph{de Bruijn graph}
data structure is commonly used~\cite{bruijn1946combinatorial}.
Often, however, a \emph{reference genome} (or in the case of RNA, 
\emph{transcriptome}) is available for the organism being sequenced; the
establishment of a human reference genome was indeed the purpose of the Human
Genome Project.

@@

Techniques for reference-based read mapping typically rely on algorithmic 
approaches such as the Burrows-Wheeler transform (BWT)~\cite{berger2013computational}, 
which provides efficient string 
compression through a reversible transformation, while the FM-index data 
structure~\cite{berger2013computational} is a compressed substring index, 
based on the BWT, which provides efficient storage as well as fast search.
BWA (Burrows-Wheeler Aligner)~\cite{berger2013computational} uses the BWT, while the 
Bowtie2~\cite{berger2013computational} mapper further relies on the FM-index for 
efficient mapping of NGS reads.
The Genome Multitool (GEM) mapper~\cite{marco2012gem} also uses an FM-index 
coupled with dynamic programming in a compressed representation of the 
reference genome, in order to prune the search space 
when mapping reads to a reference genome.
Masai~\cite{siragusa2013fast} and mrsFAST \cite{hach2010mrsfast} use an `approximate seed' approach to indexing
the space of possible matches, likewise pruning the search space; however, the
bulk of its runtime is spent on the extend phase.
State-of-the-art mapper mrsFAST-Ultra achieves improvements in efficiency based on machine architecture rather than leveraging redundancy in the data itself with near-perfect sensitivity, but
only for the case where there are no insertions and deletions (indels) \cite{hach2014mrsfast}.
Even with these approaches, read mapping remains a significant bottlenecks in
genomic research~\cite{berger2013computational}.
